---
title: "Reproducing XGBoost Models Results for Gauged and Ungauged Catchments"
author: "Hamed Sharif"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

Following results can be reproduced with this document:

  1. Spatial maps of gauged behavioral classes in dormant and growing seasons,
  
  2. Performance metrics of trained XGBoost models in dormant and growing seasons over test and training sets of gauged catchments,
  
  3. Regional performance metrics of trained XGBoost model in dormant and growing seasons, and
  
  4. Predicted behavioral classes for ungauged catchments in dormant and growing seasons.


```{r get packages and data, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,
                      warning = FALSE, dpi = 300,
                      fig.align = 'center')
options(knitr.kable.NA = '-')
library(data.table)
library(here)
library(tidyverse)
library(xgboost)
library(caret)
library(knitr)
library(countrycode)
library(maptiles)
library(tidyterra)
library(sf)
library(scales)
library(ggthemes)
library(rnaturalearth)

# Test and Training IDs for dormant and growing trained models -----------------
load(here("data", "Training_IDs.RData"))
dormant_model <- xgb.load(here("code", "models", "dormant_model"))
growing_model <- xgb.load(here("code", "models", "growing_model"))

# Gauged catchments metadata and attributes ------------------------------------
df_atts_gauged <- fread(here("data", 
                             "Gauged_Catchments_Metadata_and_Attributes.csv")) %>%
  dplyr::filter(!Catchment_Boundary_Flagged) %>%
  mutate(
    dormant_gauged_class = ordered(
      dormant_gauged_class,
      levels = c("simple", "intermediate", "complex")
    ),
    growing_gauged_class = ordered(
      growing_gauged_class,
      levels = c("simple", "intermediate", "complex")
    ),
    dormant_predicted_class = ordered(
      dormant_predicted_class,
      levels = c("simple", "intermediate", "complex")
    ),
    growing_predicted_class = ordered(
      growing_predicted_class,
      levels = c("simple", "intermediate", "complex")
    )
  )

# XGBoost features for gauged catchments ---------------------------------------
xgboost_gauged_atts <- df_atts_gauged %>%
  # renaming for easier tracking of features
  rename(CTI = `Async Index x Slope/AI`,
         `CTIW-dormant` = `RW1 x Async Index x Slope/AI-dormant`,
         `CTIW-growing` = `RW1 x Async Index x Slope/AI-growing`) %>%
  pivot_longer(cols = c(`RW1-dormant`:`RW5-growing`, `CTIW-dormant`, `CTIW-growing`),
               names_to = c("var", "period"), values_to = "val",
               names_sep = "-") %>%
  pivot_wider(names_from = var, values_from = val) %>%
  dplyr::select(c(period, GCIN, AI, RW1, RW2.5, RW5, Area, `Async Index`,
                  `Slope/AI`, CTI, CTIW, `(P-AET) x Slope`, 
                  `Median DepthToBedrock`, `Median TWI`, 
                  `First Geo Dominant Class`, `Second Geo Dominant Class`,
                  `Frequency Not Level`, `G Shape Elevation`,
                  `G Shape Slope`, `G Shape TWI`, `G Shape TRI`, 
                  `G Scale Elevation`, `G Scale Slope`, `G Scale TWI`,
                  `G Scale TRI`, Urban, Forest, Cropland, Bare))

# Regions for gauged catchments ------------------------------------------------
na <- df_atts_gauged %>%
  dplyr::filter(country %in% c("CA", "US")) %>% pull(GCIN)

sa <- df_atts_gauged %>% 
  mutate(
    country = ifelse(country == "AUS", "AU", country),
    continent = countrycode(country, origin = "iso2c",
                            destination = "region")
  ) %>%
  dplyr::filter(continent == "Latin America & Caribbean") %>% pull(GCIN)

eu <- df_atts_gauged %>% 
  mutate(
    country = ifelse(country == "AUS", "AU", country),
    continent = countrycode(country, origin = "iso2c",
                            destination = "continent")
  ) %>%
  dplyr::filter(continent == "Europe") %>% pull(GCIN)

af <- df_atts_gauged %>% 
  mutate(
    country = ifelse(country == "AUS", "AU", country),
    continent = countrycode(country, origin = "iso2c",
                            destination = "continent")
  ) %>%
  dplyr::filter(continent == "Africa") %>% pull(GCIN)

oc <- df_atts_gauged %>% 
  mutate(
    country = ifelse(country == "AUS", "AU", country),
    continent = countrycode(country, origin = "iso2c",
                            destination = "continent")
  ) %>%
  dplyr::filter(continent == "Oceania") %>% pull(GCIN)

regions_ids <- data.frame(
  region = c(
    rep("NA", length(na)), rep("SA", length(sa)), rep("EU", length(eu)),
    rep("AF", length(af)), rep("OC", length(oc))
  ),
  GCIN = c(na, sa, eu, af, oc)
)

# Ungauged catchments metadata and attributes ----------------------------------
df_atts_ungauged <- fread(here("data", 
                               "Ungauged_Catchments_Metadata_and_Attributes.csv")) %>%
  mutate(
    dormant_predicted_class = ordered(
      dormant_predicted_class,
      levels = c("simple", "intermediate", "complex")
    ),
    growing_predicted_class = ordered(
      growing_predicted_class,
      levels = c("simple", "intermediate", "complex")
    ),
    country = ifelse(country == "", "NA", country) # Namibia was read as #N/A
  )

# XGBoost features for ungauged catchments -------------------------------------
xgboost_ungauged_atts <- df_atts_ungauged %>%
  # renaming for easier tracking of features
  rename(CTI = `Async Index x Slope/AI`,
         `CTIW-dormant` = `RW1 x Async Index x Slope/AI-dormant`,
         `CTIW-growing` = `RW1 x Async Index x Slope/AI-growing`) %>%
  pivot_longer(cols = c(`RW1-dormant`:`RW5-growing`, `CTIW-dormant`, `CTIW-growing`),
               names_to = c("var", "period"), values_to = "val",
               names_sep = "-") %>%
  pivot_wider(names_from = var, values_from = val) %>%
  dplyr::select(c(period, UCIN, AI, RW1, RW2.5, RW5, Area, `Async Index`,
                  `Slope/AI`, CTI, CTIW, `(P-AET) x Slope`, 
                  `Median DepthToBedrock`, `Median TWI`, 
                  `First Geo Dominant Class`, `Second Geo Dominant Class`,
                  `Frequency Not Level`, `G Shape Elevation`,
                  `G Shape Slope`, `G Shape TWI`, `G Shape TRI`, 
                  `G Scale Elevation`, `G Scale Slope`, `G Scale TWI`,
                  `G Scale TRI`, Urban, Forest, Cropland, Bare))
```


```{r handy functions}
# function to get the performance metrics for both training set and test set
# as well as all samples
calculate_class_metrics <- function(conf_matrix, class_index) {
  true_positives <- conf_matrix[class_index, class_index]
  false_positives <- sum(conf_matrix[, class_index]) - true_positives
  false_negatives <- sum(conf_matrix[class_index, ]) - true_positives
  true_negatives <- sum(conf_matrix) - 
    (true_positives + false_positives + false_negatives)
  
  precision <- true_positives / (true_positives + false_positives)
  recall <- true_positives / (true_positives + false_negatives)
  specificity <- true_negatives / (true_negatives + false_positives)
  f1_score <- 2 * (precision * recall) / (precision + recall)
  
  list(precision = precision, recall = recall,
       specificity = specificity, f1_score = f1_score)
}

eval_metrics_exe <- function(model, dmatrix, obs_labels) {
  # Predict on the test set
  pred <- predict(model, newdata = dmatrix, reshape = TRUE)
  pred_labels <- max.col(pred) - 1
  
  conf_mat <- confusionMatrix(ordered(pred_labels, levels = c(0, 1, 2)),
                              ordered(as.numeric(obs_labels) - 1, 
                                      levels = c(0, 1, 2)))
  # calculate simple performances
  performances_simple <- calculate_class_metrics(conf_mat$table, 1)
  # calculate intermediate performances
  performances_intermediate <- calculate_class_metrics(conf_mat$table, 2)
  # calculate complex performance
  performances_complex <- calculate_class_metrics(conf_mat$table, 3)
  
  # evaluating complex vs non-complex
  testLabels_adj <- ifelse(obs_labels %in% c("simple", "intermediate"),
                           "non-complex", "complex")
  testLabels_adj <- ordered(testLabels_adj,
                            levels = c("non-complex", "complex"))
  pred_labels_adj <- ifelse(pred_labels %in% c(0, 1), 0, 1)
  conf_mat_adj <- confusionMatrix(ordered(pred_labels_adj, levels = c(0, 1)),
                                  ordered(as.numeric(testLabels_adj) - 1,
                                          levels = c(0, 1)))
  performances_noncomplex <- calculate_class_metrics(conf_mat_adj$table, 1) 
  performances_complex_adj <- calculate_class_metrics(conf_mat_adj$table, 2) 
  
  return(
    data.frame(
      behavioral_class = c("Simple", "Intermediate", "Complex", "All Non-Complex",
                           "All Complex"),
      num_samples = c(
        summary(obs_labels), summary(testLabels_adj)
      ),
      precision = c(performances_simple$precision,
                    performances_intermediate$precision,
                    performances_complex$precision,
                    performances_noncomplex$precision,
                    performances_complex_adj$precision),
      specificity = c(performances_simple$specificity,
                      performances_intermediate$specificity,
                      performances_complex$specificity,
                      performances_noncomplex$specificity,
                      performances_complex_adj$specificity)
    )
  )
}
```

# 1. Gauged Catchments

```{r}
gauged_sf <- df_atts_gauged %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(3857)
```

## 1.1. Dormant Season

### 1.1.1. Gauged Behavioral Classes

```{r, fig.width=7, fig.height=3.5, fig.cap="Gauged Catchments Behavioral Classes in Dormant Season"}
bg_basemap <- get_tiles(
  x = st_bbox(c(xmin = -180, xmax = 180, ymax = 65, ymin = -65), crs = st_crs(4326)), 
  provider = "Esri.WorldTopoMap", 
  zoom = 1, 
  crop = TRUE
)

ggplot() +
  geom_spatraster_rgb(data = bg_basemap, alpha = 0.5) +
  geom_sf(data = gauged_sf %>%
            arrange(desc(dormant_gauged_class == "complex"), 
                    desc(dormant_gauged_class == "intermediate")),
          aes(fill = dormant_gauged_class), size = 0.7, 
          stroke = 0.1, shape = 21) +
  scale_fill_manual(
    values = c(
      "simple" = alpha("#00459B", 0.7),
      "intermediate" = alpha("#78A455", 0.7),
      "complex" = alpha("#C06844", 0.7)
    )
  ) +
  labs(fill = " Gauged Behavioral Class in Dormant (-)") +
  theme_map() +
  theme(
    legend.position = "top", legend.direction = "horizontal",
    legend.title.position = "top"
  )
```


---

### 1.1.2. Training and Test Sets Performance Metrics

80% have been used for training, and 20% for test set.

Number of test set catchments: `r nrow(df_atts_gauged) - length(training_ids_dormant)`

Number of training set catchments: `r length(training_ids_dormant)`

```{r}
# Preparing Training Set
trainData <- xgboost_gauged_atts %>%
  dplyr::filter(period == "dormant") %>%
  dplyr::filter(GCIN %in% training_ids_dormant) %>%
  left_join(df_atts_gauged %>%
              dplyr::select(c(GCIN, dormant_gauged_class)),
            by = "GCIN") %>%
  dplyr::select(-c(GCIN, period)) %>%
  rename(Class = dormant_gauged_class)
trainFeatures <- trainData %>% dplyr::select(-Class)
trainLabels <- trainData$Class
dtrain <- xgb.DMatrix(data = as.matrix(trainFeatures), 
                      label = as.numeric(trainLabels) - 1)

# Preparing Test Set
testData <- xgboost_gauged_atts %>%
  dplyr::filter(period == "dormant") %>%
  dplyr::filter(!GCIN %in% training_ids_dormant) %>%
  left_join(df_atts_gauged %>%
              dplyr::select(c(GCIN, dormant_gauged_class)),
            by = "GCIN") %>%
  dplyr::select(-c(GCIN, period)) %>%
  rename(Class = dormant_gauged_class)
testFeatures <- testData %>% dplyr::select(-Class)
testLabels <- testData$Class
dtest <- xgb.DMatrix(data = as.matrix(testFeatures), 
                     label = as.numeric(testLabels) - 1)

# Get performance metrics: specificity and precision for each class
bind_rows(
  eval_metrics_exe(dormant_model, dtrain, trainLabels) %>%
    mutate(Set = "Training") %>%
    relocate(Set),
  eval_metrics_exe(dormant_model, dtest, testLabels) %>%
    mutate(Set = "Test") %>%
    relocate(Set)
) %>%
  mutate(
    behavioral_class = ordered(
      behavioral_class,
      levels = c("Simple", "Intermediate", "Complex", "All Non-Complex",
                 "All Complex")
    )
  ) %>%
  dplyr::select(-c(num_samples)) %>%
  kable(
    col.names = c("Set", "Class", "Precision", "Specificity"),
    digits = 2,
    align = "lccc"
  )
```

---


### 1.1.3. Regional Performance Metrics

```{r}
xgdata_regional <- xgboost_gauged_atts %>%
  dplyr::filter(period == "dormant") %>%
  left_join(df_atts_gauged %>%
              dplyr::select(c(GCIN, dormant_gauged_class)),
            by = "GCIN") %>%
  dplyr::select(-c(period)) %>%
  rename(Class = dormant_gauged_class)

result_regions_df <- data.frame()
for (n in c("NA", "SA", "EU", "AF", "OC")) {
  ids <- regions_ids$GCIN[which(regions_ids$region == n)]
  dtest_iter <- xgb.DMatrix(
    data = as.matrix(
      xgdata_regional[which(xgdata_regional$GCIN %in% ids), ] %>%
        dplyr::select(-c(GCIN, Class))
    ), 
    label = as.numeric(
      xgdata_regional[which(xgdata_regional$GCIN %in% ids), ] %>%
        pull(Class)
    ) - 1
  )
  labels_iter <- xgdata_regional[which(xgdata_regional$GCIN %in% ids), ] %>%
    pull(Class)
  result_regions_df <- rbind(
    result_regions_df,
    eval_metrics_exe(dormant_model, dtest_iter, labels_iter) %>%
      mutate(Set = n) %>%
      relocate(Set)
  )
}

result_regions_df %>%
  rename(Region = Set) %>%
    mutate(
      Region = case_when(
        Region == "NA" ~ "Canada and the US",
        Region == "SA" ~ "Latin America & Caribbean",
        Region == "EU" ~ "Europe",
        Region == "AF" ~ "Africa",
        Region == "OC" ~ "Oceania"
      )
    ) %>%
  kable(
    col.names = c("Region", "Class", "Number of Catchments",
                  "Precision", "Specificity"),
    digits = 2,
    align = "lccc"
  )
```


---

## 1.2. Growing Season

### 1.2.1. Gauged Behavioral Classes

```{r, fig.width=7, fig.height=3.5, fig.cap="Gauged Catchments Behavioral Classes in Growing Season"}
ggplot() +
  geom_spatraster_rgb(data = bg_basemap, alpha = 0.5) +
  geom_sf(data = gauged_sf %>%
            drop_na(growing_gauged_class) %>%
            arrange(desc(growing_gauged_class == "complex"), 
                    desc(growing_gauged_class == "intermediate")),
          aes(fill = growing_gauged_class), size = 0.7, 
          stroke = 0.1, shape = 21) +
  scale_fill_manual(
    values = c(
      "simple" = alpha("#00459B", 0.7),
      "intermediate" = alpha("#78A455", 0.7),
      "complex" = alpha("#C06844", 0.7)
    )
  ) +
  labs(fill = " Gauged Behavioral Class in Growing (-)") +
  theme_map() +
  theme(
    legend.position = "top", legend.direction = "horizontal",
    legend.title.position = "top"
  )
```


---

### 1.2.2. Training and Test Sets Performance Metrics

80% have been used for training, and 20% for test set.

Number of test set catchments: `r nrow(df_atts_gauged %>% dplyr::filter(!is.na(growing_gauged_class))) - length(training_ids_growing)`

Number of training set catchments: `r length(training_ids_growing)`

```{r}
# Preparing Training Set
trainData <- xgboost_gauged_atts %>%
  dplyr::filter(period == "growing") %>%
  dplyr::filter(GCIN %in% training_ids_growing) %>%
  left_join(df_atts_gauged %>%
              dplyr::select(c(GCIN, growing_gauged_class)),
            by = "GCIN") %>%
  dplyr::select(-c(GCIN, period)) %>%
  rename(Class = growing_gauged_class) %>%
  drop_na(Class)
trainFeatures <- trainData %>% dplyr::select(-Class)
trainLabels <- trainData$Class
dtrain <- xgb.DMatrix(data = as.matrix(trainFeatures), 
                      label = as.numeric(trainLabels) - 1)

# Preparing Test Set
testData <- xgboost_gauged_atts %>%
  dplyr::filter(period == "growing") %>%
  dplyr::filter(!GCIN %in% training_ids_growing) %>%
  left_join(df_atts_gauged %>%
              dplyr::select(c(GCIN, growing_gauged_class)),
            by = "GCIN") %>%
  dplyr::select(-c(GCIN, period)) %>%
  rename(Class = growing_gauged_class) %>%
  drop_na(Class)
testFeatures <- testData %>% dplyr::select(-Class)
testLabels <- testData$Class
dtest <- xgb.DMatrix(data = as.matrix(testFeatures), 
                     label = as.numeric(testLabels) - 1)

# Get performance metrics: specificity and precision for each class
bind_rows(
  eval_metrics_exe(growing_model, dtrain, trainLabels) %>%
    mutate(Set = "Training") %>%
    relocate(Set),
  eval_metrics_exe(growing_model, dtest, testLabels) %>%
    mutate(Set = "Test") %>%
    relocate(Set)
) %>%
  mutate(
    behavioral_class = ordered(
      behavioral_class,
      levels = c("Simple", "Intermediate", "Complex", "All Non-Complex",
                 "All Complex")
    )
  ) %>%
  dplyr::select(-c(num_samples)) %>%
  kable(
    col.names = c("Set", "Class", "Precision", "Specificity"),
    digits = 2,
    align = "lccc"
  )
```

---


### 1.2.3. Regional Performance Metrics

```{r}
xgdata_regional <- xgboost_gauged_atts %>%
  dplyr::filter(period == "growing") %>%
  left_join(df_atts_gauged %>%
              dplyr::select(c(GCIN, growing_gauged_class)),
            by = "GCIN") %>%
  dplyr::select(-c(period)) %>%
  rename(Class = growing_gauged_class) %>%
  drop_na(Class)

result_regions_df <- data.frame()
  for (n in c("NA", "SA", "EU", "AF", "OC")) {
    ids <- regions_ids$GCIN[which(regions_ids$region == n)]
    dtest_iter <- xgb.DMatrix(
      data = as.matrix(
        xgdata_regional[which(xgdata_regional$GCIN %in% ids), ] %>%
          dplyr::select(-c(GCIN, Class))
      ), 
      label = as.numeric(
        xgdata_regional[which(xgdata_regional$GCIN %in% ids), ] %>%
          pull(Class)
      ) - 1
    )
    labels_iter <- xgdata_regional[which(xgdata_regional$GCIN %in% ids), ] %>%
      pull(Class)
    result_regions_df <- rbind(
      result_regions_df,
      eval_metrics_exe(growing_model, dtest_iter, labels_iter) %>%
        mutate(Set = n) %>%
      relocate(Set)
    )
  }

result_regions_df %>%
  rename(Region = Set) %>%
    mutate(
      Region = case_when(
        Region == "NA" ~ "Canada and the US",
        Region == "SA" ~ "Latin America & Caribbean",
        Region == "EU" ~ "Europe",
        Region == "AF" ~ "Africa",
        Region == "OC" ~ "Oceania"
      )
    ) %>%
  kable(
    col.names = c("Region", "Class", "Number of Catchments",
                  "Precision", "Specificity"),
    digits = 2,
    align = "lccc"
  )
```


---


# 2. Ungauged Catchments

Here, static map of predicted behavioral classes for ungauged catchments are reproduced. An interactive map of predicted behavioral classes for ungauged catchments were also developed that can be accessed via the following URL: https://h-sharif.github.io/stormflow-behavior/

## 2.1. Predicted Behavioral Classes in Dormant Seasons

```{r, fig.width=7, fig.height=3.5, fig.cap="Distribution of Ungauged Catchments Predicted Behavioral Classes in Dormant Season"}
df_atts_ungauged %>%
  count(dormant_predicted_class) %>%
  mutate(pct = n / sum(n)) %>%
  ggplot(aes(x = dormant_predicted_class, y = pct)) +
  geom_col() +
  geom_text(
    aes(label = n),
    vjust = -0.5,
    size = 3.5
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.1)),
                     limits = c(0, 1), breaks = seq(0, 1, 0.2),
                     minor_breaks = seq(0, 1, 0.1)) +
  labs(
    x = "Predicted Behavioral Class in Dormant",
    y = "Percentage (%)",
    subtitle = "Total Number of Ungauged Catchments in Dormant Season: 77,585",
    fill = ""
  ) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal")
```

---

```{r, fig.width=7, fig.height=3.5, fig.cap="Distribution of Ungauged Catchments Area for each Predicted Behavioral Class in Dormant Season"}
df_atts_ungauged %>%
  group_by(dormant_predicted_class) %>%
  summarise(total_area = sum(Area) / 1e6,
            .groups = "drop") %>%
  mutate(area_pct = total_area / sum(total_area)) %>%
  ggplot(aes(x = dormant_predicted_class, y = area_pct)) +
  geom_col() +
  geom_text(
    aes(label = paste(round(total_area), "sq.km")),
    vjust = -0.5,
    size = 3.5
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.1)),
                     limits = c(0, 1), breaks = seq(0, 1, 0.2),
                     minor_breaks = seq(0, 1, 0.1)) +
  labs(
    x = "Predicted Behavioral Class in Dormant",
    y = "Percentage of Total Processed Area (%)",
    subtitle = expression("Total Area of All Ungauged Catchments: 139 x "*10^6~"("*km^2*")"),
    fill = ""
  ) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal")
```


---

### 2.1.1. Canada (BC) and Salish Sea Region

```{r, fig.width=5, fig.height=6.5, fig.cap="Predicted Behavioral Classes for Ungauged Catchments in Canada (BC) and Salish Sea Region in Dormant Season"}
sf_obj <- df_atts_ungauged %>%
  dplyr::filter(`Polygon Source` == "BCUB") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(3857)

bg_basemap <- get_tiles(
  x = st_bbox(sf_obj %>% st_buffer(100000)), 
  provider = "Esri.WorldTopoMap", 
  zoom = 5, 
  crop = TRUE
)

ggplot() +
  geom_spatraster_rgb(data = bg_basemap, alpha = 0.7) +
  geom_sf(data = sf_obj %>%
            arrange(desc(dormant_predicted_class == "complex"), 
                    desc(dormant_predicted_class == "intermediate")),
          aes(color = dormant_predicted_class), shape = 16, size = 0.3) +
  scale_color_manual(
    values = c(
      "simple" = alpha("#00459B", 0.2),
      "intermediate" = alpha("#78A455", 0.2),
      "complex" = alpha("#C06844", 0.2)
    )
  ) +
  labs(color = "Predicted Behavioral Class in Dormant (-)",
       subtitle = "Ungauged Catchments in Canada (BC) and Salish Sea Region") +
  theme_map() +
  theme(
    legend.position = "bottom", legend.direction = "horizontal",
    legend.title.position = "top"
  ) +
  guides(color = guide_legend(override.aes = list(size = 1.5, alpha = 1)))
```


---

```{r}
sf_obj %>%
  st_drop_geometry() %>%
  group_by(dormant_predicted_class) %>%
  summarise(
    n = n(),
    area = round(sum(Area)/1e6, digits = 3),
    .groups ="drop"
  ) %>%
  kable(
    col.names = c("Predicted Behavioral Class", "Number of Catchments", "$Area\\,(10^6\\,km^2)$"),
    caption = "Canada (BC) and Salish Sea Region in Dormant Season",
    align = "lcc"
  )
```


---


### 2.1.2. Europe

```{r, fig.width=8, fig.height=6.5, fig.cap="Predicted Behavioral Classes for Ungauged Catchments in Europe in Dormant Season"}
sf_obj <- df_atts_ungauged %>%
  mutate(continent = countrycode(country, origin = "iso2c", destination = "continent"))  %>%
  dplyr::filter(continent == "Europe") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

sf_bbox <- st_bbox(sf_obj)

europe_borders <- ne_countries(scale = "medium", continent = "Europe", returnclass = "sf")

europe_labels <- europe_borders %>%
  # Filter to map area so text doesn't crowd out the edges
  filter(admin %in% c("United Kingdom", "France", "Germany", "Spain", "Italy", 
                      "Austria", "Belgium", "Netherlands", "Switzerland", "Denmark")) %>%
  mutate(iso_a2 = ifelse(iso_a2 == "-99", "FR", iso_a2)) # fixing France code

ggplot() +
  geom_sf(data = europe_borders, fill = alpha("grey80", 0.2), color = "grey30",
          linewidth = 0.1) +
  geom_sf(data = sf_obj %>%
            arrange(desc(dormant_predicted_class == "complex"), 
                    desc(dormant_predicted_class == "intermediate")),
          aes(fill = dormant_predicted_class), shape = 21, size = 0.7, stroke = 0.03) +
  geom_sf_text(data = europe_labels, aes(label = iso_a2), 
               color = "grey40", fontface = "bold", size = 2, alpha = 0.8) +
  scale_fill_manual(
    values = c(
      "simple" = alpha("#00459B", 0.8),
      "intermediate" = alpha("#78A455", 0.8),
      "complex" = alpha("#C06844", 0.8)
    )
  ) +
  labs(fill = "Predicted Behavioral Class in Dormant (-)",
       subtitle = "Ungauged Catchments in Europe") +
  theme_map() +
  theme(
    legend.position = "bottom", legend.direction = "horizontal",
    legend.title.position = "top"
  ) +
  guides(fill = guide_legend(override.aes = list(size = 1.5, alpha = 1))) +
  coord_sf(xlim = c(sf_bbox["xmin"], sf_bbox["xmax"]), 
           ylim = c(sf_bbox["ymin"], sf_bbox["ymax"]), 
           expand = TRUE)
```


---

```{r}
sf_obj %>%
  st_drop_geometry() %>%
  group_by(dormant_predicted_class) %>%
  summarise(
    n = n(),
    area = round(sum(Area)/1e6, digits = 3),
    .groups ="drop"
  ) %>%
  kable(
    col.names = c("Predicted Behavioral Class", "Number of Catchments", "$Area\\,(10^6\\,km^2)$"),
    caption = "Europe in Dormant Season",
    align = "lcc"
  )
```


---


### 2.1.3. Latin America

```{r, fig.width=5, fig.height=6, fig.cap="Predicted Behavioral Classes for Ungauged Catchments in Latin America in Dormant Season"}
sf_obj <- df_atts_ungauged %>%
  mutate(continent = countrycode(country, origin = "iso2c", destination = "continent"))  %>%
  dplyr::filter(continent == "Americas", !country %in% c("CA", "US")) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

sf_bbox <- st_bbox(sf_obj)

sa_borders <- ne_countries(scale = "medium", returnclass = "sf")

sa_labels <- sa_borders %>%
  filter(admin %in% c("Mexico", "Brazil", "Argentina", "Colombia", "Peru", 
                      "Venezuela", "Chile", "Ecuador", "Bolivia", "Paraguay", 
                      "Uruguay", "Guyana", "Suriname", "Panama", "Costa Rica"))

ggplot() +
  geom_sf(data = sa_borders, fill = alpha("grey80", 0.2), color = "grey30",
          linewidth = 0.1) +
  geom_sf(data = sf_obj %>%
            arrange(desc(dormant_predicted_class == "complex"), 
                    desc(dormant_predicted_class == "intermediate")),
          aes(fill = dormant_predicted_class), shape = 21, size = 0.6, stroke = 0.03) +
  geom_sf_text(data = sa_labels, aes(label = iso_a2), 
               color = "grey40", fontface = "bold", size = 2, alpha = 0.8) +
  scale_fill_manual(
    values = c(
      "simple" = alpha("#00459B", 0.8),
      "intermediate" = alpha("#78A455", 0.8),
      "complex" = alpha("#C06844", 0.8)
    )
  ) +
  labs(fill = "Predicted Behavioral Class in Dormant (-)",
       subtitle = "Ungauged Catchments in Latin America") +
  theme_map() +
  theme(
    legend.position = "bottom", legend.direction = "horizontal",
    legend.title.position = "top"
  ) +
  guides(fill = guide_legend(override.aes = list(size = 1.5, alpha = 1))) +
  coord_sf(xlim = c(sf_bbox["xmin"], sf_bbox["xmax"]), 
           ylim = c(sf_bbox["ymin"], sf_bbox["ymax"]), 
           expand = TRUE)
```

---

```{r}
sf_obj %>%
  st_drop_geometry() %>%
  group_by(dormant_predicted_class) %>%
  summarise(
    n = n(),
    area = round(sum(Area)/1e6, digits = 3),
    .groups ="drop"
  ) %>%
  kable(
    col.names = c("Predicted Behavioral Class", "Number of Catchments", "$Area\\,(10^6\\,km^2)$"),
    caption = "Latin America in Dormant Season",
    align = "lcc"
  )
```


---

### 2.1.4. Australia

```{r, fig.width=7, fig.height=5, fig.cap="Predicted Behavioral Classes for Ungauged Catchments in Austrila and New Zealand in Dormant Season"}
nz_borders <- ne_countries(scale = "medium", country = "New Zealand", returnclass = "sf")
aus_states <- ne_states(country = "Australia", returnclass = "sf")

sf_obj <- df_atts_ungauged %>%
  mutate(continent = countrycode(country, origin = "iso2c", destination = "continent"))  %>%
  dplyr::filter(continent == "Oceania") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

sf_bbox <- st_bbox(sf_obj)

ggplot() +
  geom_sf(data = nz_borders, fill = alpha("grey80", 0.2), color = "grey30", linewidth = 0.1) +
  geom_sf(data = aus_states, fill = alpha("grey80", 0.2), color = "grey30", linewidth = 0.1) +
  geom_sf(data = sf_obj %>%
            arrange(desc(dormant_predicted_class == "complex"), 
                    desc(dormant_predicted_class == "intermediate")),
          aes(fill = dormant_predicted_class), shape = 21, size = 0.8, stroke = 0.03) +
  scale_fill_manual(
    values = c(
      "simple" = alpha("#00459B", 0.8),
      "intermediate" = alpha("#78A455", 0.8),
      "complex" = alpha("#C06844", 0.8)
    )
  ) +
  labs(fill = "Predicted Behavioral Class in Dormant (-)",
       subtitle = "Ungauged Catchments in Oceania") +
  theme_map() +
  theme(
    legend.position = "bottom", legend.direction = "horizontal",
    legend.title.position = "top"
  ) +
  guides(fill = guide_legend(override.aes = list(size = 1.5, alpha = 1))) +
  coord_sf(xlim = c(sf_bbox["xmin"], sf_bbox["xmax"]), 
           ylim = c(sf_bbox["ymin"], sf_bbox["ymax"]), 
           expand = TRUE)
```


---

```{r}
sf_obj %>%
  st_drop_geometry() %>%
  group_by(dormant_predicted_class) %>%
  summarise(
    n = n(),
    area = round(sum(Area)/1e6, digits = 3),
    .groups ="drop"
  ) %>%
  kable(
    col.names = c("Predicted Behavioral Class", "Number of Catchments", "$Area\\,(10^6\\,km^2)$"),
    caption = "Oceania in Dormant Season",
    align = "lcc"
  )
```

---

## 2.2. Predicted Behavioral Classes in Growing Seasons

```{r, fig.width=7, fig.height=3.5, fig.cap="Distribution of Ungauged Catchments Predicted Behavioral Classes in Growing Season"}
df_atts_ungauged %>%
  drop_na(growing_predicted_class) %>%  # growing dormancy info were not available for two catchments in AUS
  count(growing_predicted_class) %>%
  mutate(pct = n / sum(n)) %>%
  ggplot(aes(x = growing_predicted_class, y = pct)) +
  geom_col() +
  geom_text(
    aes(label = n),
    vjust = -0.5,
    size = 3.5
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.1)),
                     limits = c(0, 1), breaks = seq(0, 1, 0.2),
                     minor_breaks = seq(0, 1, 0.1)) +
  labs(
    x = "Predicted Behavioral Class in Growing",
    y = "Percentage (%)",
    subtitle = "Total Number of Ungauged Catchments in Growing Season: 77,583",
    fill = ""
  ) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal")
```

---

```{r, fig.width=7, fig.height=3.5, fig.cap="Distribution of Ungauged Catchments Area for each Predicted Behavioral Class in Growing Season"}
df_atts_ungauged %>%
  drop_na(growing_predicted_class) %>%  # growing dormancy info were not available for two catchments in AUS
  group_by(growing_predicted_class) %>%
  summarise(total_area = sum(Area) / 1e6,
            .groups = "drop") %>%
  mutate(area_pct = total_area / sum(total_area)) %>%
  ggplot(aes(x = growing_predicted_class, y = area_pct)) +
  geom_col() +
  geom_text(
    aes(label = paste(round(total_area), "sq.km")),
    vjust = -0.5,
    size = 3.5
  ) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.1)),
                     limits = c(0, 1), breaks = seq(0, 1, 0.2),
                     minor_breaks = seq(0, 1, 0.1)) +
  labs(
    x = "Predicted Behavioral Class in Growing",
    y = "Percentage of Total Processed Area (%)",
    subtitle = expression("Total Area of All Ungauged Catchments: 139 x "*10^6~"("*km^2*")"),
    fill = ""
  ) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal")
```

---

### 2.2.1. Canada (BC) and Salish Sea Region

```{r, fig.width=5, fig.height=6.5, fig.cap="Predicted Behavioral Classes for Ungauged Catchments in Canada (BC) and Salish Sea Region in Growing Season"}
sf_obj <- df_atts_ungauged %>%
  dplyr::filter(`Polygon Source` == "BCUB") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(3857)

bg_basemap <- get_tiles(
  x = st_bbox(sf_obj %>% st_buffer(100000)), 
  provider = "Esri.WorldTopoMap", 
  zoom = 5, 
  crop = TRUE
)

ggplot() +
  geom_spatraster_rgb(data = bg_basemap, alpha = 0.7) +
  geom_sf(data = sf_obj %>%
            arrange(desc(growing_predicted_class == "complex"), 
                    desc(growing_predicted_class == "intermediate")),
          aes(color = growing_predicted_class), shape = 16, size = 0.3) +
  scale_color_manual(
    values = c(
      "simple" = alpha("#00459B", 0.2),
      "intermediate" = alpha("#78A455", 0.2),
      "complex" = alpha("#C06844", 0.2)
    )
  ) +
  labs(color = "Predicted Behavioral Class in Growing (-)",
       subtitle = "Ungauged Catchments in Canada (BC) and Salish Sea Region") +
  theme_map() +
  theme(
    legend.position = "bottom", legend.direction = "horizontal",
    legend.title.position = "top"
  ) +
  guides(color = guide_legend(override.aes = list(size = 1.5, alpha = 1)))
```


---

```{r}
sf_obj %>%
  st_drop_geometry() %>%
  group_by(growing_predicted_class) %>%
  summarise(
    n = n(),
    area = round(sum(Area)/1e6, digits = 3),
    .groups ="drop"
  ) %>%
  kable(
    col.names = c("Predicted Behavioral Class", "Number of Catchments", "$Area\\,(10^6\\,km^2)$"),
    caption = "Canada (BC) and Salish Sea Region in Growing Season",
    align = "lcc"
  )
```

---

### 2.2.2. Europe

```{r, fig.width=8, fig.height=6.5, fig.cap="Predicted Behavioral Classes for Ungauged Catchments in Europe in Growing Season"}
sf_obj <- df_atts_ungauged %>%
  mutate(continent = countrycode(country, origin = "iso2c", destination = "continent"))  %>%
  dplyr::filter(continent == "Europe") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

sf_bbox <- st_bbox(sf_obj)

europe_borders <- ne_countries(scale = "medium", continent = "Europe", returnclass = "sf")

europe_labels <- europe_borders %>%
  # Filter to map area so text doesn't crowd out the edges
  filter(admin %in% c("United Kingdom", "France", "Germany", "Spain", "Italy", 
                      "Austria", "Belgium", "Netherlands", "Switzerland", "Denmark")) %>%
  mutate(iso_a2 = ifelse(iso_a2 == "-99", "FR", iso_a2)) # fixing France code

ggplot() +
  geom_sf(data = europe_borders, fill = alpha("grey80", 0.2), color = "grey30",
          linewidth = 0.1) +
  geom_sf(data = sf_obj %>%
            arrange(desc(growing_predicted_class == "complex"), 
                    desc(growing_predicted_class == "intermediate")),
          aes(fill = growing_predicted_class), shape = 21, size = 0.7, stroke = 0.03) +
  geom_sf_text(data = europe_labels, aes(label = iso_a2), 
               color = "grey40", fontface = "bold", size = 2, alpha = 0.8) +
  scale_fill_manual(
    values = c(
      "simple" = alpha("#00459B", 0.8),
      "intermediate" = alpha("#78A455", 0.8),
      "complex" = alpha("#C06844", 0.8)
    )
  ) +
  labs(fill = "Predicted Behavioral Class in Growing (-)",
       subtitle = "Ungauged Catchments in Europe") +
  theme_map() +
  theme(
    legend.position = "bottom", legend.direction = "horizontal",
    legend.title.position = "top"
  ) +
  guides(fill = guide_legend(override.aes = list(size = 1.5, alpha = 1))) +
  coord_sf(xlim = c(sf_bbox["xmin"], sf_bbox["xmax"]), 
           ylim = c(sf_bbox["ymin"], sf_bbox["ymax"]), 
           expand = TRUE)
```

---


```{r}
sf_obj %>%
  st_drop_geometry() %>%
  group_by(growing_predicted_class) %>%
  summarise(
    n = n(),
    area = round(sum(Area)/1e6, digits = 3),
    .groups ="drop"
  ) %>%
  kable(
    col.names = c("Predicted Behavioral Class", "Number of Catchments", "$Area\\,(10^6\\,km^2)$"),
    caption = "Europe in Growing Season",
    align = "lcc"
  )
```


---

### 2.2.3. Latin America

```{r, fig.width=5, fig.height=6, fig.cap="Predicted Behavioral Classes for Ungauged Catchments in Latin America in Growing Season"}
sf_obj <- df_atts_ungauged %>%
  mutate(continent = countrycode(country, origin = "iso2c", destination = "continent"))  %>%
  dplyr::filter(continent == "Americas", !country %in% c("CA", "US")) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

sf_bbox <- st_bbox(sf_obj)

sa_borders <- ne_countries(scale = "medium", returnclass = "sf")

sa_labels <- sa_borders %>%
  filter(admin %in% c("Mexico", "Brazil", "Argentina", "Colombia", "Peru", 
                      "Venezuela", "Chile", "Ecuador", "Bolivia", "Paraguay", 
                      "Uruguay", "Guyana", "Suriname", "Panama", "Costa Rica"))

ggplot() +
  geom_sf(data = sa_borders, fill = alpha("grey80", 0.2), color = "grey30",
          linewidth = 0.1) +
  geom_sf(data = sf_obj %>%
            arrange(desc(growing_predicted_class == "complex"), 
                    desc(growing_predicted_class == "intermediate")),
          aes(fill = growing_predicted_class), shape = 21, size = 0.6, stroke = 0.03) +
  geom_sf_text(data = sa_labels, aes(label = iso_a2), 
               color = "grey40", fontface = "bold", size = 2, alpha = 0.8) +
  scale_fill_manual(
    values = c(
      "simple" = alpha("#00459B", 0.8),
      "intermediate" = alpha("#78A455", 0.8),
      "complex" = alpha("#C06844", 0.8)
    )
  ) +
  labs(fill = "Predicted Behavioral Class in Growing (-)",
       subtitle = "Ungauged Catchments in Latin America") +
  theme_map() +
  theme(
    legend.position = "bottom", legend.direction = "horizontal",
    legend.title.position = "top"
  ) +
  guides(fill = guide_legend(override.aes = list(size = 1.5, alpha = 1))) +
  coord_sf(xlim = c(sf_bbox["xmin"], sf_bbox["xmax"]), 
           ylim = c(sf_bbox["ymin"], sf_bbox["ymax"]), 
           expand = TRUE)
```

---


```{r}
sf_obj %>%
  st_drop_geometry() %>%
  group_by(growing_predicted_class) %>%
  summarise(
    n = n(),
    area = round(sum(Area)/1e6, digits = 3),
    .groups ="drop"
  ) %>%
  kable(
    col.names = c("Predicted Behavioral Class", "Number of Catchments", "$Area\\,(10^6\\,km^2)$"),
    caption = "Latin America in Growing Season",
    align = "lcc"
  )
```


---

### 2.2.4. Australia

```{r, fig.width=7, fig.height=5, fig.cap="Predicted Behavioral Classes for Ungauged Catchments in Austrila and New Zealand in Growing Season"}
nz_borders <- ne_countries(scale = "medium", country = "New Zealand", returnclass = "sf")
aus_states <- ne_states(country = "Australia", returnclass = "sf")

sf_obj <- df_atts_ungauged %>%
  drop_na(growing_predicted_class) %>% # growing dormancy info were not available for two catchments in AUS
  mutate(continent = countrycode(country, origin = "iso2c", destination = "continent"))  %>%
  dplyr::filter(continent == "Oceania") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

sf_bbox <- st_bbox(sf_obj)

ggplot() +
  geom_sf(data = nz_borders, fill = alpha("grey80", 0.2), color = "grey30", linewidth = 0.1) +
  geom_sf(data = aus_states, fill = alpha("grey80", 0.2), color = "grey30", linewidth = 0.1) +
  geom_sf(data = sf_obj %>%
            arrange(desc(growing_predicted_class == "complex"), 
                    desc(growing_predicted_class == "intermediate")),
          aes(fill = growing_predicted_class), shape = 21, size = 0.8, stroke = 0.03) +
  scale_fill_manual(
    values = c(
      "simple" = alpha("#00459B", 0.8),
      "intermediate" = alpha("#78A455", 0.8),
      "complex" = alpha("#C06844", 0.8)
    )
  ) +
  labs(fill = "Predicted Behavioral Class in Growing (-)",
       subtitle = "Ungauged Catchments in Oceania") +
  theme_map() +
  theme(
    legend.position = "bottom", legend.direction = "horizontal",
    legend.title.position = "top"
  ) +
  guides(fill = guide_legend(override.aes = list(size = 1.5, alpha = 1))) +
  coord_sf(xlim = c(sf_bbox["xmin"], sf_bbox["xmax"]), 
           ylim = c(sf_bbox["ymin"], sf_bbox["ymax"]), 
           expand = TRUE)
```

---


```{r}
sf_obj %>%
  st_drop_geometry() %>%
  group_by(growing_predicted_class) %>%
  summarise(
    n = n(),
    area = round(sum(Area)/1e6, digits = 3),
    .groups ="drop"
  ) %>%
  kable(
    col.names = c("Predicted Behavioral Class", "Number of Catchments", "$Area\\,(10^6\\,km^2)$"),
    caption = "Oceania in Growing Season",
    align = "lcc"
  )
```


---
